---
output:
  pdf_document: default
  html_document: default
---
##############################################################
# 
# Load libraries
#
##############################################################
```{r}
# Clear all variables in the workplace
rm(list=ls())

# Load the forecasting package
library(fpp2)
library(pacman)

p_load(tidyverse, lubridate, stringr, scales, cowplot, dygraphs, xts, imputeTS, forecast, zoo, fma, expsmooth, fpp2, doParallel)
```
# Load Data
```{r}
path <- "D:/An I5/TSA/Mini Project/Code/household_power_consumption.txt"

# Import the data
data <- read_delim(path, ";", escape_double = FALSE, trim_ws = TRUE)

```
#############################################
# Data preprocessing
#############################################
```{r}
# Convert the Date column to Date type
data <- data %>%
  mutate(Date = dmy(Date))  # Assuming the format is Day-Month-Year

# Convert Time column to proper time format (if needed)
data <- data %>%
  mutate(Time = hms(Time))  # Requires lubridate for parsing time
```

# Handling Invalid and Missing Values
## Replace invalid by NA 
## Replace NA by mean
```{r}
# Check for problematic rows in all columns
invalid_rows <- data %>%
  filter(!str_detect(Global_active_power, "^\\d+(\\.\\d+)?$") |
           !str_detect(Global_reactive_power, "^\\d+(\\.\\d+)?$") |
           !str_detect(Voltage, "^\\d+(\\.\\d+)?$") |
           !str_detect(Global_intensity, "^\\d+(\\.\\d+)?$") |
           !str_detect(Sub_metering_1, "^\\d+(\\.\\d+)?$") |
           !str_detect(Sub_metering_2, "^\\d+(\\.\\d+)?$") |
           !str_detect(Sub_metering_3, "^\\d+(\\.\\d+)?$"))

print(invalid_rows)

# Replace invalid values with NA for numeric columns
data <- data %>%
  mutate_at(vars(Global_active_power, Global_reactive_power, Voltage, 
                 Global_intensity, Sub_metering_1, Sub_metering_2, Sub_metering_3), 
            ~ ifelse(str_detect(., "^\\d+(\\.\\d+)?$"), as.numeric(.), NA))

# Replace NA with column mean
data <- data %>%
  mutate_at(vars(Global_active_power, Global_reactive_power, Voltage, 
                 Global_intensity, Sub_metering_1, Sub_metering_2, Sub_metering_3), 
            ~ replace(., is.na(.), mean(., na.rm = TRUE)))

```
# Convert character columns to numeric
```{r}
# Combine Date and Time into a single Datetime column
data <- data %>%
  mutate(Datetime = as.POSIXct(paste(Date, Time)))

# Convert character columns to numeric
data <- data %>%
  mutate_at(vars(Global_active_power, Global_reactive_power, Voltage, 
                 Global_intensity, Sub_metering_1, Sub_metering_2), 
            ~ as.numeric(.))
```

# View Data
```{r}
# Display the structure of the dataset
print(str(data))

# Display a summary of the dataset
print(summary(data))

# Count missing values in each column
print(colSums(is.na(data)))
```

###################################################
# Explore Global Active Power (GAP)
###################################################
```{r}
# select datetime and Global
gap_data <- data %>%
  select(Date, Time, Datetime, Global_active_power)

# Display structure and summary
str(gap_data)
summary(gap_data)
colSums(is.na(gap_data))
```
# Visualization
```{r}
# Load necessary libraries
library(ggplot2)

# Histogram of Global_active_power
ggplot(gap_data, aes(x = Global_active_power)) +
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  labs(title = "Distribution of Global Active Power", x = "Global Active Power (kW)", y = "Frequency")

# Boxplot of Global_active_power to check for outliers
ggplot(gap_data, aes(y = Global_active_power)) +
  geom_boxplot(fill = "blue", color = "black") +
  labs(title = "Boxplot of Global Active Power", y = "Global Active Power (kW)")

# Time series plot of Global_active_power over time
ggplot(gap_data, aes(x = Datetime, y = Global_active_power)) +
  geom_line(color = "blue") +
  labs(title = "Global Active Power over Time", x = "Datetime", y = "Global Active Power (kW)")
```
# Remove outlier
```{r}
# Function to remove rows with outliers in a column
remove_outliers_from_df <- function(df, column_name) {
  # Calculate the IQR
  IQR_value <- IQR(df[[column_name]], na.rm = TRUE)
  Q1 <- quantile(df[[column_name]], 0.25, na.rm = TRUE)
  Q3 <- quantile(df[[column_name]], 0.75, na.rm = TRUE)
  
  # Define lower and upper bounds for outliers
  lower_bound <- Q1 - 1.5 * IQR_value
  upper_bound <- Q3 + 1.5 * IQR_value
  
  # Filter the data frame to keep only rows within bounds
  df <- df[df[[column_name]] >= lower_bound & df[[column_name]] <= upper_bound, ]
  
  return(df)
}



# gap_data1 <- gap_data
# Apply the function to the entire data frame for the specified column
gap_data <- remove_outliers_from_df(gap_data, "Global_active_power")
```
# Visualize again
```{r visualize-gap}
# Load necessary libraries
library(ggplot2)

# Histogram for Global_active_power
ggplot(gap_data, aes(x = Global_active_power)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Global Active Power", x = "Global Active Power", y = "Frequency") +
  theme_minimal()

# Box plot for Global_active_power
ggplot(gap_data, aes(y = Global_active_power)) +
  geom_boxplot(fill = "lightgreen", color = "black") +
  labs(title = "Box Plot of Global Active Power", y = "Global Active Power") +
  theme_minimal()

# Time series plot for Global_active_power over time
ggplot(gap_data, aes(x = Datetime, y = Global_active_power)) +
  geom_line(color = "blue") +
  labs(title = "Time Series of Global Active Power", x = "Datetime", y = "Global Active Power") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r aggregate-gap}
library(dplyr)
library(ggplot2)
library(lubridate)

# Assuming `gap_data` has a "Date" column in Date format and "Global_active_power" column

# 1. Aggregate data to daily level
daily_gap_data <- gap_data %>%
  group_by(Date) %>%
  summarise(Daily_Avg = mean(Global_active_power, na.rm = TRUE))

# Convert to a daily time series
daily_ts <- ts(daily_gap_data$Daily_Avg, start = c(2007, 1), frequency = 365)

# 2. Aggregate data to weekly level
weekly_data <- daily_gap_data %>%
  mutate(Week = floor_date(Date, unit = "week")) %>%
  group_by(Week) %>%
  summarise(Weekly_Avg = mean(Daily_Avg, na.rm = TRUE))

# Convert to a weekly time series
weekly_ts <- ts(weekly_data$Weekly_Avg, start = c(2007, 1), frequency = 52)

# 3. Aggregate data to monthly level
monthly_data <- daily_gap_data %>%
  mutate(Month = floor_date(Date, unit = "month")) %>%
  group_by(Month) %>%
  summarise(Monthly_Avg = mean(Daily_Avg, na.rm = TRUE))

# Convert to a monthly time series
monthly_ts <- ts(monthly_data$Monthly_Avg, start = c(2007, 1), frequency = 12)

# 4. Aggregate data to quarterly level
quarterly_data <- daily_gap_data %>%
  mutate(Quarter = paste0(year(Date), " Q", quarter(Date))) %>%
  group_by(Quarter) %>%
  summarise(Quarterly_Avg = mean(Daily_Avg, na.rm = TRUE))

# Convert to a quarterly time series
quarterly_ts <- ts(quarterly_data$Quarterly_Avg, start = c(2007, 1), frequency = 4)

# 5. Plotting examples
# Daily Plot
ggplot(daily_gap_data, aes(x = Date, y = Daily_Avg)) +
  geom_line(color = "blue") +
  labs(title = "Daily Average Global Active Power", x = "Date", y = "Average Global Active Power (kW)")

# Weekly Plot
ggplot(weekly_data, aes(x = Week, y = Weekly_Avg)) +
  geom_line(color = "green") +
  labs(title = "Weekly Average Global Active Power", x = "Week", y = "Average Global Active Power (kW)")

# Monthly Plot
ggplot(monthly_data, aes(x = Month, y = Monthly_Avg)) +
  geom_line(color = "red") +
  labs(title = "Monthly Average Global Active Power", x = "Month", y = "Average Global Active Power (kW)")


```



###################################################################
# Daily time series decomposed (time plot, trend, season, error)
###################################################################
```{r}
library(dplyr)
Y <- monthly_ts

# Decompose the quarterly time series
decomposed_Y <- decompose(Y, type = "additive")

# Plot the decomposition
plot(decomposed_Y)

```
# Box plot
```{r}
boxplot(Y, main = "Boxplot of Global Active Power", ylab = "Global Active Power")
Q1 <- quantile(Y, 0.25)
Q3 <- quantile(Y, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
outliers <- Y[Y < lower_bound | Y > upper_bound]
print(outliers)
```
```{r}
DY <- diff(Y)

autoplot(DY) +
  ggtitle(" Time Plot")
  ylab("KW")
  

ggseasonplot(DY) +
  ggtitle("Season Plot")
  ylab("KW")
  
ggsubseriesplot(DY)
```
```{r}
if (!require(tseries)) install.packages("tseries")
library(tseries)
adf_test <- adf.test(DY)
print(adf_test)
```
```{r}
# we can use Y instead of DY in auto.arima
fit_arima <- auto.arima(Y, stepwise = FALSE, approximation = FALSE, trace=TRUE)
print(summary(fit_arima))
checkresiduals(fit_arima)

# ACF of residuals (manually if you want more control)
acf(residuals(fit_arima), main = "ACF of ARIMA Residuals")

# Histogram of residuals
hist(residuals(fit_arima), main = "Histogram of ARIMA Residuals", xlab = "Residuals", breaks = 30)

# QQ-plot of residuals
qqnorm(residuals(fit_arima))
qqline(residuals(fit_arima), col = "red")

# Ljung-Box test to check for autocorrelation in residuals
Box.test(residuals(fit_arima), lag = 20, type = "Ljung-Box")
### arima(2,1,3) SD^2 = 0.09049 SD=0.3008156
```
# Daily Forecasting
```{r}
fcst <- forecast(fit_arima, h=63)
autoplot(fcst, include=1095)
```
```{r}
# Fit ARIMA model
# fit_arima <- auto.arima(Y, stepwise = FALSE, approximation = FALSE, trace = TRUE)

# Print ARIMA model summary
print(summary(fit_arima))

# Check residual diagnostics
checkresiduals(fit_arima)

# ACF of residuals
acf(residuals(fit_arima), main = "ACF of ARIMA Residuals")

# Histogram of residuals
hist(residuals(fit_arima), 
     main = "Histogram of ARIMA Residuals", 
     xlab = "Residuals", 
     breaks = 30, 
     col = "lightblue", 
     border = "black")

# QQ-plot of residuals
qqnorm(residuals(fit_arima), main = "QQ-Plot of ARIMA Residuals")
qqline(residuals(fit_arima), col = "red")

# Ljung-Box test for autocorrelation in residuals
box_test <- Box.test(residuals(fit_arima), lag = 20, type = "Ljung-Box")
print(box_test)

# Shapiro-Wilk test for normality of residuals
shapiro_test <- shapiro.test(residuals(fit_arima))
print(shapiro_test)

# Jarque-Bera test for normality of residuals
if (!require(tseries)) install.packages("tseries")
library(tseries)
jarque_bera_test <- jarque.bera.test(residuals(fit_arima))
print(jarque_bera_test)

# Runs Test for randomness of residuals
if (!require(randtests)) install.packages("randtests")
library(randtests)
runs_test <- runs.test(residuals(fit_arima))
print(runs_test)

# Calculate variance and standard deviation of residuals
residual_variance <- var(residuals(fit_arima))
residual_sd <- sqrt(residual_variance)
cat("Residual Variance:", residual_variance, "\n")
cat("Residual Standard Deviation:", residual_sd, "\n")

```
Interpretation of Diagnostic Tests:
Box-Ljung Test:

X-squared = 12.886, df = 20, p-value = 0.8822
Interpretation: The high p-value (> 0.05) suggests that there is no significant autocorrelation in the residuals. This is a good sign, as it indicates that the ARIMA model has captured the time series' patterns effectively.
Shapiro-Wilk Normality Test:

W = 0.87703, p-value = 0.0001235
Interpretation: The low p-value (< 0.05) indicates that the residuals deviate significantly from a normal distribution. This suggests that the residuals may not be normally distributed, which might affect the validity of statistical inferences based on the model.
Jarque-Bera Test:

X-squared = 44.074, df = 2, p-value = 2.688e-10
Interpretation: The extremely low p-value (< 0.05) confirms that the residuals are not normally distributed. This aligns with the result of the Shapiro-Wilk test. The non-normality may be due to skewness, kurtosis, or other distributional characteristics.
Runs Test:

Statistic = -3.5016, p-value = 0.0004626
Interpretation: The low p-value (< 0.05) indicates non-randomness in the residuals. This suggests that there may be patterns left in the residuals that the model has not captured, implying the ARIMA model may need improvement or adjustments.
Residual Variance and Standard Deviation:

Residual Variance = 0.0091
Residual Standard Deviation = 0.0956
Interpretation: The residual variance and standard deviation indicate the average dispersion of the residuals. While these values are relatively low, the presence of non-randomness and non-normality suggests further refinement of the model may be required.


